LabMate: supporting types for MATLAB

### 2024-04-10 writing/IMEKOWC2024/LabMate/paper.tex
[ writing ] fix Section 5

Coming back to the example from Section~\ref{sec:example} again, we can give both \texttt{A} and \texttt{B} more precise types, also taking their physical dimensions into account:
\begin{verbatim}
% > A :: [ i <- 12,  j <- 5 ]
      (if j == 4 then Q T else Q 1)
% > B :: [ i <- 5 ]
      (if i == 4
         then Q (ML^2T^(-4)I^(-1))
         else Q (ML^2T^(-3)I^(-1)))
\end{verbatim}
This says that the entries in the last column of \texttt{A} are quantities of dimension time $\mathsf{T}$, while all other entries are dimensionless quantities. Meanwhile the entries of \texttt{B} are quantities of dimension $\mathsf{ML}^{2}\mathsf{T}^{-3}\mathsf{I}^{-1}$ (i.e., units V) for entries in the first three rows and $\mathsf{ML}^{2}\mathsf{T}^{-4}\mathsf{I}^{-1}$ (i.e., units V/s) respectively. LabMate reports that the type of \texttt{C} should be
\begin{verbatim}
%< C :: [ 5 x 1 ] (Q (ML^2T^(-3)I^(-1)))
\end{verbatim}
that is a $5 \times 1$ matrix whose entries are all quantities of dimension $\mathsf{ML}^{2}\mathsf{T}^{-3}\mathsf{I}^{-1}$. Another possible way to use LabMate is to instead declare that this is the expected type for \texttt{C}, and to leave the type of one of the entries in \texttt{B} as an unknown:
\begin{verbatim}
% > B :: [ Q TODO; Q TODO; ?aTy;
           Q TODO; Q TODO ]
\end{verbatim}
We can then ask LabMate to solve the \remph{meta variable} \texttt{?aTy}, to which LabMate responds that we must have \todo{\texttt{?aTy} = fill in}. This way the types of LabMate can be used to validate and experiment with the dimensional consistency of the quantities involved.

@@

Thanks to NPL colleague Nick Fletcher for his help and guidance providing the example from Section \ref{sec:example}. Thanks also to NPL colleagues Louise Wright and Ian Smith for reviewing this paper.

>>>>

Coming back to the example from Section~\ref{sec:example} again, we can give both \texttt{A} and \texttt{B} more precise types, also taking their physical dimensions into account:
\begin{verbatim}
%> A :: [ i <- 12,  j <- 5 ]
          (if j == 4 then Q T else Q 1)

%> b :: [ 5 x 1 ] (Q (ML^2T^(-3)I^(-1)))
\end{verbatim}
This says that the entries in the last column of \texttt{A} are quantities of dimension time $\mathsf{T}$, while all other entries are dimensionless quantities. Meanwhile all the entries of \texttt{b} are quantities of dimension $\mathsf{ML}^{2}\mathsf{T}^{-3}\mathsf{I}^{-1}$. LabMate reports that the type of \texttt{x} should be
\begin{verbatim}
%< x :: [ i <- 5 ]
        (if i == 4
           then Q (ML^2T^(-4)I^(-1))
           else Q (ML^2T^(-3)I^(-1)))
\end{verbatim}
that is a $5 \times 1$ matrix whose entries are all quantities of dimension $\mathsf{ML}^{2}\mathsf{T}^{-3}\mathsf{I}^{-1}$ (i.e., units V) for entries in the first four rows and $\mathsf{ML}^{2}\mathsf{T}^{-4}\mathsf{I}^{-1}$ (i.e., units V/s) in the final row.
%
Another possible way to use LabMate is to instead declare that this is the expected type for \texttt{x}, and to leave the type of some of the entries in \texttt{b} as an unknown
\begin{verbatim}
%> b :: [ i <- 5 ]
        (if i == 4
           then ?bTy
           else Q (ML^2T^(-3)I^(-1)))
\end{verbatim}
We can then ask LabMate to solve the \remph{meta variable} \texttt{?bTy}, to which LabMate responds that we must also have \texttt{?aTy} = $\mathsf{Q}\,(\mathsf{ML}^{2}\mathsf{T}^{-3}\mathsf{I}^{-1})$. This way the types of LabMate can be used to validate and experiment with the dimensional consistency of the quantities involved.

@@

Thanks to NPL colleague Nick Fletcher for his help and guidance providing the example from Section \ref{sec:example}. Thanks also to NPL colleagues Louise Wright and Ian Smith for reviewing this paper, and to Professor Neil Ghani for his support.
